---
title: Housing Price Predictions
author: Gabe Mednick
date: '2021-05-18'
slug: AmesHousing
categories: []
tags: []
subtitle: ''
summary: 'Compare logistic regression to k-nearest neighbors clustering using the tidymodels framework for machine learning'
authors: []
lastmod: ''
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

library(tidyverse)
library(tidymodels)
library(AmesHousing)
library(leaflet)
theme_set(theme_light())
```

## This post is a work in progress

In this post, we will explore the Ames housing dataset and use it to create a model for price prediction. The dataset, available  in the `AmesHousing` package, contains 81 features for 3000 home sales between 2006-2010. It is a clean version of the dataset, so no cleaning is needed. However, we will need to do some data engineering later on.  

```{r}
ames_df <- make_ames() %>% 
  janitor::clean_names()

# Here are a few functions that I used to get an initial feel for the dataset
#ames_df %>% count(neighborhood)
#summary(ames_df)
#View(ames_df)
#dim(ames_df)
```

Sale price is the feature of interest or response variable. Let's take a closer by plotting it as a distribution. Even before we visualize sale price as a histogram, we can imagine that the distribution might be something like a bell-shape with a skewed tale owing to a few uber-expensive mansions. In this case, we will want to apply a log transformation to the sale price before modeling. This will be addressed when we get to the data engineering step.

```{r}
library(scales)
library(patchwork)
sale_price_dist <- ggplot(ames_df, aes(x = sale_price)) + 
  geom_histogram(bins = 50) +
  geom_vline(lty = 2, color = 'red', xintercept = mean(ames_df$sale_price)) +
  scale_x_continuous(labels = dollar_format()) +
  labs(x = 'sale price')

sale_price_log_dist <- ggplot(ames_df, aes(x = sale_price)) + 
  geom_histogram(bins = 50) +
  geom_vline(lty = 2, color = 'red', xintercept = mean(ames_df$sale_price)) +
  #scale_x_continuous(labels = dollar_format())  +
  scale_x_log10() +
  labs(x = 'log(sale price)')

sale_price_dist | sale_price_log_dist 
```

How has the sale price changed with time? The plot below shows the average sale price plotted over the four years of available data. It gives us a general feel for the housing market over the four year period that the data was recorded. Grouping the data by neighborhood and plotting it was too busy. Instead, the data was grouped by year and the average sale price was calculated. 

```{r}
p <- ames_df %>% select(neighborhood, year_sold, sale_price) %>% 
  group_by(year_sold) %>% 
  mutate(avg_sale_price = mean(sale_price)) %>% 
  arrange(year_sold) %>% 
  ggplot(aes(year_sold, avg_sale_price)) +
  geom_line() +
  theme(legend.position = 'none') +
  scale_y_continuous(labels = comma)

plotly::ggplotly(p)
```


The dataset includes longitude and latitude coordinates for each house. It is reasonable to assume that sale price might have a strong correlation with neighborhood. In the map below, houses (circles) are colored by neighborhood and the size of the circle is weighted by sale price. In order to better visualize the individual circles, `dplyr::slice_sample()` can be used to randomly sample a limited number of houses--we are looking at 600 in the Leaflet map shown below.


```{r}
library(RColorBrewer)
pal <- colorFactor(
  palette = 'Dark2',
  domain = ames_df$neighborhood
)

library(RColorBrewer)
pal <- colorFactor(
  palette = 'Dark2',
  domain = ames_df$neighborhood
)
ames_random_samp <- ames_df %>% 
  slice_sample(n = 600)

ames_random_samp %>% group_by(neighborhood) %>% leaflet(data = .,
        options = leafletOptions(zoomControl = FALSE,
                                 minZoom = 12, maxZoom = 12, dragging = FALSE)) %>% 
  addTiles() %>%  
  addCircles(~longitude, ~latitude, color = ~pal(neighborhood), radius = 1, weight = ~sale_price*2e-5, opacity = 1)

```

Now that we have a better feel for the data, let's split it into a training and test set using the `rsample` package. We will also use `vfold_cv()` to resample the training set. By default `vfold_cv` will split the data into 10 folds (v) where each fold is split into a training and test set. 

```{r}
library(recipes)

set.seed(518)
ames_split <- initial_split(ames_df, prop = 0.8, strata = "sale_price") 

ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

Data engineering with the `recipes` package, provides an arsenal of data transformations that may be necessary and/or beneficial to the predictive power of our trained machine learning model. 

```{r}
library(recipes) #loaded with the tidymodels package

ames_rec <- 
  recipe(sale_price ~ ., data = ames_train) %>%
  step_log(sale_price, base = 10) %>% 
  step_other(neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal()) 

ames_lm <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")

lm_wkfl <- 
  workflow() %>% 
  add_model(ames_lm) %>% 
  add_recipe(ames_rec)

ames_wkfl_fit <- lm_wkfl %>%
  last_fit(split = ames_split)

ames_wkfl_fit %>%
  collect_metrics()

ames_preds <- ames_wkfl_fit %>%
  collect_predictions()

# ames_test_rec <- ames_rec %>%
#   prep() %>% 
#   bake(new_data = ames_test) 
#   
# ames_results <- ames_test_rec %>% 
#   select(sale_price) %>% 
#   bind_cols(ames_preds)

ggplot(ames_preds, aes(x = sale_price, y = .pred)) +
  geom_point() +
  geom_abline(color = 'blue', linetype = 2) +
  coord_obs_pred() +
  labs(title = 'R-Squared Plot',
       y = 'Predicted Sale Price (log)',
       x = 'Actual Sale Price (log)')
```


